COVID-19 infection and attributable mortality in UK long term care facilities: Cohort study using active surveillance and electronic records (March-June 2020)

### 2020-11-27 manuscript/manuscript.Rmd
Revisions version 1.4

---
title: "Covid-19 infection and attributable mortality in UK long term care facilities: Cohort study using active surveillance and electronic records (March-June 2020)"
bibliography: references.bib
csl: sage-vancouver.csl
output:
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
  word_document:
    toc: yes
    reference_docx: template.docx
link-citations: yes
always_allow_html: yes
---

Peter Dutey Magni,$^{12}$ Haydn Williams,$^{3}$ Arnoupe Jhass,$^{4}$ Greta Rait,$^4$ Harry Hemingway,$^{125}$ Andrew Hayward,$^{6}$ Laura Shallcross$^{12*}$

@@

$^6$ Institute of Epidemiology & Health Care, University College London, London, UK

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
knit_output_pdf <- F
if(knit_output_pdf){
  options(knitr.table.format = "pdf")
} else {
  options(knitr.table.format = "pandoc")
}

source("../.Rprofile")
devtools::load_all("..")
load_data()
library(dplyr)
library(lubridate)
library(kableExtra)
library(captioner)
library(survival)
library(survminer)
library(ggplot2)
library(finalfit)
# override the finalfit cox PH modelling which defaults to Efron ties
# using Breslow ties instead
# coxphmulti <- function (.data, dependent, explanatory) {
#     requireNamespace("survival")
#     coxph(as.formula(paste0(dependent, "~", paste(explanatory, 
#         collapse = "+"))), data = .data, ties = "breslow")
# }
# assignInNamespace("coxphmulti",coxphmulti,ns="finalfit")

@@

residents_overview <- tline %>% 
  filter(rday==1) %>%
  mutate(ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  group_by(resident_id, home_code, ageg, gender, 
           admission_type2, home_type, user_type2, status) %>% 
  summarise(resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed),
            symptomatic = max(-susceptible_symptomatic+1, na.rm = T),
            confirmed = max(-susceptible_confirmed+1, na.rm = T),
            start_date = min(date, na.rm = T),
            stop_date = max(date, na.rm = T),
            start_int = as.numeric(min(date, na.rm = T) - date_min),
            stop_int = as.numeric(max(date, na.rm = T)- date_min)
    ) %>% 
  ungroup() %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Uninfected",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Uninfected", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed"))) %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, staff_to_bed_ratio_g5, carer_to_bed_ratio, bed_tot_category,
                   beds,  pct_occupancy, no_bedrooms, uk_imd_england_quintile,
                   occ_bedroom_ratio, occ_bedroom_ratio_g5, occ_bedroom_ratio_g6)) %>% 
  left_join(distinct(select(incidents, resident_id, starts_with("covid_first"))))

home_first_case <- bind_rows(
  new_cases %>% 
    dplyr::filter(tidyr::replace_na(new_deaths_hospital, 0) > 0 |
                    tidyr::replace_na(new_deaths_home, 0)>0) %>% 
    dplyr::group_by(home_code) %>% 
    dplyr::summarise(first_case_date = min(date),
                     source = "CT"),
  residents_overview %>% 
    dplyr::filter(confirmed == 1) %>% 
    dplyr::group_by(home_code) %>% 
    dplyr::summarise(first_case_date = min(covid_first_confirmed, na.rm = T), source="DAT")
) %>% 
  dplyr::group_by(home_code) %>% 
  dplyr::summarise(first_case_date = min(first_case_date)) 
  # tidyr::crossing(tibble(outbreak = c(0, 1)))

residents_episodes <- tline %>% 
  filter(rday==1)  %>% 
  left_join(home_first_case) %>%
  mutate(outbreak = dplyr::case_when(
    date >= first_case_date ~ 1L,
    date < first_case_date ~ 0L,
    TRUE ~ 0L
  ))  %>% 
  group_by(resident_id, home_code, dob, gender, 
           status, home_type, user_type2,
           susceptible_symptomatic, 
           susceptible_confirmed,
           outbreak) %>% 
  summarise(entry_date = min(date),
            exit_date = max(date+1),
            resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(date1 = as.numeric( entry_date - date_min),
         date2 = as.numeric(exit_date - date_min),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1),
         date2 = if_else(date1 == date2, date2 + 1, date2)) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Uninfected",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Uninfected", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed")),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95))   %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, staff_to_bed_ratio_g5, carer_to_bed_ratio, bed_tot_category, 
                   beds, pct_occupancy, no_bedrooms, uk_imd_england_quintile,
                   occ_bedroom_ratio, occ_bedroom_ratio_g5, occ_bedroom_ratio_g6)) %>% 
  group_by(resident_id) %>% 
  mutate(death_status = if_else(
    date1 == max(date1),
    as.integer(status == "Deceased"),
    0L
  )) %>% 
  ungroup()

@@

km_ct_residents_deaths <- residents_ct %>% 
  group_by(date) %>% 
  summarise(
    deaths = sum(new_deaths),
    occupancy = sum(occupancy)
  ) %>% 
  km_ct_estimator(df = ., t = date, d = deaths, 
                  pop = occupancy, subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog),
    event = "Covid-19-related death"
  )

@@

incidence_CT_residents <- table_nums(
  name = "incidence_CT_residents",
  caption =  paste0("Cumulative incidence and rate of SARS-CoV-2 infections in residents according to FSHCG manager counts (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
supp_epi_CT_oubreak_only <- table_nums(
  name = "incidence_CT_staff",
  caption =  paste0("Cumulative incidence and rate of SARS-CoV-2 infections among staff according to FSHCG manager counts (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Characteristics of FSHCG private bed residents by care home type, sex, age, region and status on study exit (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_CFR_age_sex <- table_nums(
  name = "tab_CFR_age_sex",
  caption =  paste0("All-cause case-fatality rates by age and sex among private residents (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  "; ", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Risk factors for confirmed infection in private residents: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))
tab_CoxPH_death <- table_nums(
  name = "tab_CoxPH_death",
  caption =  paste0("Risk factors for all-cause mortality in private residents of care homes with and without Covid-19 outbreaks: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_episodes$resident_id)),
  ", ", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_attributable_deaths <- table_nums(
  name = "tab_attributable_deaths",
  caption =  paste0("Model-based estimates of attributable death in private residents of care homes with and without Covid-19 outbreaks (n=", 
  formatbm(dplyr::n_distinct(residents_episodes$resident_id)),
  ", ", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))

@@

fig_study_diagram <- fig_nums(
  name = "fig_study_diagram",
  caption = paste0("Study overview: location of FSHCG care homes and diagram of data sources")
)

fig_KM_CT_plots <- fig_nums(
  name = "fig_KM_CT_plots",
  caption = paste0("Kaplan-Meier estimates of the cumulative incidence of symptomatic cases, confirmed infections and COVID-related deaths in (A) residents (n=",   formatbm(desc_residents$CT_total_residents),
  ") and (B) care home staff (n=",   formatbm(desc_staff$total_staff),
  ") according to FSHCG outbreak surveillance counts (", format(date_min_tallies, "%e %b %Y"), "-", format(date_max_tallies, "%e %b %Y"), ")"))

@@

**Background** Rates of Covid-19 infection have declined in many countries, but outbreaks persist in care home residents who are at high risk of severe outcomes. Epidemiological data from care homes are scarce. We used population-level active surveillance to estimate incidence of, and risk factors for Covid-19, and attributable mortality in care home residents.

**Methods:** Cohort study using individual-level electronic health records from 8,713 residents and daily counts of infection for `r formatbm(desc_residents$CT_total_residents)` residents and `r formatbm(desc_staff$total_staff)` staff across `r dplyr::n_distinct(residents$home_code)` UK care homes. We modelled risk factors for infection and mortality using Cox proportional hazards and estimated attributable fractions.

**Findings:** `r formatbm(desc_residents$CT_total_symptomatic)` residents developed Covid-19 symptoms (`r tbrounding(desc_residents$CT_p_symptomatic)`% [95% confidence interval: `r tbrounding(desc_residents$CT_p_symptomatic_lower)`%; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`%]), while `r formatbm(desc_residents$CT_total_confirmed)` residents (`r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%]) and `r formatbm(desc_staff$CT_confirmed_total)` staff (`r tbrounding(desc_staff$CT_confirmed_p)`% [`r tbrounding(desc_staff$CT_confirmed_p_lower)`%; `r tbrounding(desc_staff$CT_confirmed_p_upper)`%]) had laboratory confirmed infections. Confirmed infection incidence in residents and staff respectively was `r tbrounding(desc_residents$CT_IR_confirmed)` [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`] and `r tbrounding(desc_staff$CT_confirmed_IR)` [`r tbrounding(desc_staff$CT_confirmed_IR_lower)`; `r tbrounding(desc_staff$CT_confirmed_IR_upper)`] per 100,000 person-days. `r length(desc_residents_outbreak_only$outbreak_homes)`/`r desc_residents$total_homes` (`r tbrounding(length(desc_residents_outbreak_only$outbreak_homes)/desc_residents$total_homes*100)`%) care homes had at least one Covid-19 infection or death. Lower staffing ratios and higher occupancy rates were independent risk factors for infection.

`r formatbm(desc_residents$DAT_total_deaths)` all-cause deaths occurred in `r formatbm(desc_residents$DAT_total_residents)` (`r tbrounding(desc_residents$DAT_p_death)`% [`r tbrounding(desc_residents$DAT_p_death_lower)`%; `r tbrounding(desc_residents$DAT_p_death_upper)`%]) residents. `r desc_residents$DAT_CFR$y` deaths occured in `r desc_residents$DAT_CFR$n` residents with confirmed infection (case-fatality rate: `r tbrounding(desc_residents$DAT_CFR_rate)`% [`r tbrounding(desc_residents$DAT_CFR_rate_lower)`%; `r tbrounding(desc_residents$DAT_CFR_rate_upper)`%]). Mortality in residents with no direct evidence of infection was 96% higher in care homes with at least one Covid-19 infection or death compared to care homes without (adjusted HR 1.9 [1.7; 2.2]). 33% of all-cause deaths were attributable to Covid-19 either through confirmed infection, or excess mortality.

**Interpretation:** 1 in 5 residents had symptoms of infection during the pandemic, but many cases were not tested. Higher occupancy and lower staffing levels increase infection risk. Disease control measures should integrate active surveillance and testing with fundamental changes in staffing and care home occupancy to protect staff and residents from infection.

**Funding:** Economic and Social Research Council [ES/V003887/1].

@@

Globally the number of cases of Covid-19 continues to increase, but cases in Europe have declined since April 2020,[@ECDC2020] following the introduction of lockdown measures.[@Hale2020] Although the incidence of infection in the general population in England is low (0.04%),[@cisbulletin2020] new infections persist, with substantially higher rates of infection reported in both care homes and hospitals.[@PHE2020] This raises the possibility that these settings represent a reservoir for transmission of infection back to the community.

In the UK, there are an estimated 400,000 residents living in approximately 11,000 homes for the elderly.[@LaingBuisson2019] Care home residents are particularly vulnerable to Covid-19 due to their advanced age and high prevalence of comorbidity,[@Zhou2020] and their frequent exposure to infection through close contact with staff members, other residents and contaminated surfaces in the care home setting. At the peak of the pandemic, the number of deaths in care home residents was three-fold higher than the equivalent period in in 2019.[@onsDR2020] Care home staff also have higher aged-standardised rates of Covid-19 related mortality compared to other occupations.[@onsDR2020occ] National statistics suggests two-thirds of excess deaths recorded in care home residents in the last 6 months involved Covid-19,[@onsDR2020] but this is likely to be an underestimate because many residents were not tested. Understanding the proportion of excess deaths that can be directly and indirectly attributed to Covid-19 infection is important, to fully assess the impact of the pandemic on care homes.

The development of public health strategies to protect the public, residents and staff from Covid-19 requires knowledge of the burden of and risk factors for infection in care home residents and staff, linked to outcomes. However, there is no syndromic surveillance for infection in care homes in England, and widespread one-off testing for SARS-CoV-2 using reverse transcriptase polymerase chain reaction (RT-PCR) was not established for care home staff and residents until 11 May 2020.[@dhscTesting2020] Prior to this, testing was only available for care home residents or staff who were admitted to hospital, or as part of Public Health England’s (PHE) outbreak investigations which permitted a maximum of five tests per care home. Consequently national estimates of incidence and prevalence will substantially underestimate the burden of infection in care home residents and staff.

In the absence of cohort studies or active surveillance, outbreak investigations provide the most reliable estimates of the burden of infection and case-fatality.[@McMichael2020; @Dora2020] An estimated 44% of English care homes have had at least one outbreak, with a living systematic review[@Salcher-Konrad2020] reporting substantial variation in cumulative incidence of infection (0%-72%) and case fatality (0-34%) in care home residents. A major limitation of outbreak investigations is that follow-up is usually less than 30 days,[@Salcher-Konrad2020] and investigations are only conducted in settings in which outbreaks have been detected.  Understanding the proportion of care homes with undetected cases is crucial for policy decisions around the frequency of and justification for regular testing in care homes.

To our knowledge, there are no studies which have employed population-level active surveillance in care homes throughout this pandemic to measure outcomes of both suspected and confirmed infection in residents and staff. We analysed electronic health records from the Four Seasons Health Care Group, one of the UK’s largest independent provider of residential and nursing care, with the aim of identifying strategies to protect care home staff and residents from future waves of infection. Our objectives were to estimate incidence of and risk factors for infection and mortality in the following groups: (A) residents with no evidence of infection; (B) symptomatic residents; (C) asymptomatic residents with confirmed infection; and (D) symptomatic residents with confirmed infection. We also estimated mortality attributable to Covid-19.

@@

Staff or residents living and working in care homes for the elderly between `r format(date_min, "%e %B")` and `r format(date_max, "%e %B %Y")`, which were run by the Four Seasons Healthcare Group (FSHCG) were eligible for study inclusion. The FSHCG provides a combination of residential and nursing care (for residents with medical conditions), which is primarily state funded. Most residents are permanent, but a small proportion receive temporary (respite) care.

>>>>

---
title: "COVID-19 infection and attributable mortality in UK long term care facilities: Cohort study using active surveillance and electronic records (March-June 2020)"
bibliography: references.bib
csl: sage-vancouver.csl
output:
  word_document:
    toc: no
    reference_docx: template.docx
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
link-citations: yes
always_allow_html: yes
---

Peter Dutey Magni, Researcher$^{12}$ Haydn Williams, Data manager$^{3}$ Arnoupe Jhass, General Practitioner$^{4}$ Greta Rait, Professor$^4$ Fabiana Lorencatto, Senior researcher$^5$ Harry Hemingway,$^{126}$ Andrew Hayward,$^{7}$ Laura Shallcross$^{12*}$

@@

$^6$ Centre for Behaviour Change, University College London, London, UK

$^7$ Institute of Epidemiology & Health Care, University College London, London, UK

@@

Peter Dutey-Magni p.dutey-magni@ucl.ac.uk

Haydn Williams haydn.williams@fhscgroup.com

Arnoupe Jhass a.jhass@ucl.ac.uk

Greta Rait g.rait@ucl.ac.uk

Lorencatto, Fabiana f.lorencatto@ucl.ac.uk

Harry Hemingway h.hemingway@ucl.ac.uk

Andrew Hayward a.hayward@ucl.ac.uk

Laura Shallcross l.shallcross@ucl.ac.uk

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
knit_output_pdf <- F
if(knit_output_pdf){
  options(knitr.table.format = "pdf")
} else {
  options(knitr.table.format = "pandoc")
}
source("../.Rprofile")
devtools::load_all("..")
load_data()
library(dplyr)
library(lubridate)
library(kableExtra)
library(captioner)
library(survival)
library(survminer)
library(ggplot2)
library(finalfit)
# override the finalfit cox PH modelling which defaults to Efron ties
# using Breslow ties instead
# coxphmulti <- function (.data, dependent, explanatory) {
#     requireNamespace("survival")
#     coxph(as.formula(paste0(dependent, "~", paste(explanatory, 
#         collapse = "+"))), data = .data, ties = "breslow")
# }
# assignInNamespace("coxphmulti",coxphmulti,ns="finalfit")

@@

residents_overview <- tline %>% 
  filter(rday==1) %>%
  mutate(ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  group_by(resident_id, home_code, ageg, gender, 
           admission_type2, home_type, user_type2, status) %>% 
  summarise(resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed),
            symptomatic = max(-susceptible_symptomatic+1, na.rm = T),
            confirmed = max(-susceptible_confirmed+1, na.rm = T),
            start_date = min(date, na.rm = T),
            stop_date = max(date, na.rm = T),
            start_int = as.numeric(min(date, na.rm = T) - date_min),
            stop_int = as.numeric(max(date, na.rm = T)- date_min)
    ) %>% 
  ungroup() %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Uninfected",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(
    covid_status = factor(
      covid_status, 
      levels = c("Uninfected", 
                 "Symptomatic (not confirmed)",
                 "Asymptomatic confirmed", 
                 "Symptomatic confirmed"))) %>% 
  left_join(
    select(reference_homes, home_code, country, 
           region_UK, staff_to_bed_ratio,
           staff_to_bed_ratio_g5, carer_to_bed_ratio,
           bed_tot_category, beds,  pct_occupancy,
           mean_occupancy, no_bedrooms, staff_total,
           uk_imd_england_quintile,
           occ_bedroom_ratio, occ_bedroom_ratio_g5,
           occ_bedroom_ratio_g6)) %>% 
  left_join(distinct(select(incidents, 
                            resident_id,
                            starts_with("covid_first"))))

home_first_case <- bind_rows(
  new_cases %>% 
    dplyr::filter(tidyr::replace_na(new_deaths_hospital, 0) > 0 |
                    tidyr::replace_na(new_deaths_home, 0)>0) %>% 
    dplyr::group_by(home_code) %>% 
    dplyr::summarise(first_case_date = min(date),
                     source = "CT"),
  residents_overview %>% 
    dplyr::filter(confirmed == 1) %>% 
    dplyr::group_by(home_code) %>% 
    dplyr::summarise(
      first_case_date = min(covid_first_confirmed, 
                            na.rm = T), 
      source="DAT")
  ) %>% 
  dplyr::group_by(home_code) %>% 
  dplyr::summarise(first_case_date = min(first_case_date))

residents_episodes <- tline %>% 
  filter(rday==1)  %>% 
  left_join(home_first_case) %>%
  mutate(outbreak = dplyr::case_when(
    date >= first_case_date ~ 1L,
    date < first_case_date ~ 0L,
    TRUE ~ 0L
  ))  %>% 
  group_by(resident_id, home_code, dob, gender, 
           status, home_type, user_type2,
           susceptible_symptomatic, 
           susceptible_confirmed,
           outbreak) %>% 
  summarise(entry_date = min(date),
            exit_date = max(date+1),
            resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(date1 = as.numeric( entry_date - date_min),
         date2 = as.numeric(exit_date - date_min),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1),
         date2 = if_else(date1 == date2, date2 + 1, date2)) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Uninfected",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(
    covid_status = factor(
      covid_status, 
      levels = c("Uninfected", "Symptomatic (not confirmed)",
                 "Asymptomatic confirmed", "Symptomatic confirmed")),
    ageg = cut_age_denary(compute_age(dob, date_min), 
                          age.min = 75, age.max = 95))   %>% 
  left_join(
    select(reference_homes, home_code, country, region_UK,
           staff_to_bed_ratio, staff_to_bed_ratio_g5, 
           carer_to_bed_ratio, bed_tot_category, 
           beds, pct_occupancy, no_bedrooms, uk_imd_england_quintile,
           occ_bedroom_ratio, occ_bedroom_ratio_g5, occ_bedroom_ratio_g6)) %>% 
  group_by(resident_id) %>% 
  mutate(death_status = if_else(
    date1 == max(date1),
    as.integer(status == "Deceased"),
    0L
  )) %>% ungroup()

residents_episodes <- mutate(
  residents_episodes, 
  gender = as.factor(gender),
  uk_imd_england_quintile = factor(
    uk_imd_england_quintile,
    levels = c('3', '1 - least deprived', '2',
               '4', '5 - most deprived')),
  bed_to_staff_ratio = 1/staff_to_bed_ratio
)

@@

km_ct_residents_deaths <- residents_ct %>% 
  group_by(date) %>% 
  summarise(
    deaths = sum(new_deaths),
    occupancy = sum(occupancy)
  ) %>% 
  km_ct_estimator(df = ., t = date, d = deaths, 
                  pop = occupancy, subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog),
    event = "COVID-19-related death"
  )

@@

incidence_CT_residents <- table_nums(
  name = "incidence_CT_residents",
  caption =  paste0("Cumulative incidence and rate of SARS-CoV-2 infections in residents according to FSHCG aggregate data (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
supp_epi_CT_oubreak_only <- table_nums(
  name = "incidence_CT_staff",
  caption =  paste0("Cumulative incidence and rate of SARS-CoV-2 infections among staff according to FSHCG aggregate data (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Characteristics of FSHCG private bed residents by type of LTCF, sex, age, region and status on study exit (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_CFR_age_sex <- table_nums(
  name = "tab_CFR_age_sex",
  caption =  paste0("All-cause case-fatality rates by age and sex among private residents (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  "; ", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Risk factors for confirmed infection in private residents: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))
tab_CoxPH_death <- table_nums(
  name = "tab_CoxPH_death",
  caption =  paste0("Risk factors for all-cause mortality in private residents of LTCFs with and without COVID-19 outbreaks: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_episodes$resident_id)),
  ", ", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_attributable_deaths <- table_nums(
  name = "tab_attributable_deaths",
  caption =  paste0("Model-based estimates of attributable death in private residents of LTCFs with and without COVID-19 outbreaks (n=", 
  formatbm(dplyr::n_distinct(residents_episodes$resident_id)),
  ", ", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))

@@

fig_study_diagram <- fig_nums(
  name = "fig_study_diagram",
  caption = paste0("Study overview: location of FSHCG LTCFs and diagram of data sources")
)

fig_KM_CT_plots <- fig_nums(
  name = "fig_KM_CT_plots",
  caption = paste0("Kaplan-Meier estimates of the cumulative incidence of symptomatic cases, confirmed infections and COVID-19 related deaths in (A) residents (n=",   formatbm(desc_residents$CT_total_residents),
  ") and (B) staff (n=",   formatbm(desc_staff$total_staff),
  ") according to FSHCG aggregate data (", format(date_min_tallies, "%e %b %Y"), "-", format(date_max_tallies, "%e %b %Y"), ")"))

@@

**Background**: Information on the transmission and outcomes of COVID-19 infection in staff and residents in Long term care facilities (LTCFs) in the UK is scarce. This study aimed to estimate the incidence of and risk factors for suspected and confirmed COVID-19 infection and mortality during the first wave of the pandemic.

**Methods and Findings**: 179 UK LTCFs were included in a cohort study using population-level active surveillance and linked electronic health records. Participants included 9,339 residents and 11,604 staff. We estimated the incidence of suspected and confirmed infection and mortality in staff and residents using daily tallies of infection (aggregate data) and used individual-level electronic health records from 8,713 residents to model risk factors for confirmed infection and mortality using Cox proportional hazards. We estimated mortality attributable to COVID-19 by comparing deaths in care homes with and without outbreaks.

`r formatbm(desc_residents$CT_total_symptomatic)`/`r formatbm(desc_residents$CT_total_residents)` residents developed COVID-19 symptoms (`r tbrounding(desc_residents$CT_p_symptomatic)`% [95% confidence interval: `r tbrounding(desc_residents$CT_p_symptomatic_lower)`%; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`%]), while `r formatbm(desc_residents$CT_total_confirmed)` residents (`r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%]) and `r formatbm(desc_staff$CT_confirmed_total)` staff (`r tbrounding(desc_staff$CT_confirmed_p)`% [`r tbrounding(desc_staff$CT_confirmed_p_lower)`%; `r tbrounding(desc_staff$CT_confirmed_p_upper)`%]) had laboratory confirmed infections. The rate of incidence of confirmed infection was `r tbrounding(desc_residents$CT_IR_confirmed)` [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`] and `r tbrounding(desc_staff$CT_confirmed_IR)` [`r tbrounding(desc_staff$CT_confirmed_IR_lower)`; `r tbrounding(desc_staff$CT_confirmed_IR_upper)`] per 100,000 person-days in residents and staff respectively. `r length(desc_residents_outbreak_only$outbreak_homes)`/`r desc_residents$total_homes` (`r tbrounding(length(desc_residents_outbreak_only$outbreak_homes)/desc_residents$total_homes*100)`%) LTCFs had at least one COVID-19 infection or COVID-19-related death. Lower staffing ratios and higher occupancy rates were independent risk factors for infection.

Daily tallies from LTCF managers recorded `r desc_residents$DAT_CFR$y` deaths in `r desc_residents$DAT_CFR$n` residents with confirmed infection (case-fatality rate: `r tbrounding(desc_residents$DAT_CFR_rate)`% [`r tbrounding(desc_residents$DAT_CFR_rate_lower)`%; `r tbrounding(desc_residents$DAT_CFR_rate_upper)`%]). Models estimated that 573/1,694 (33.8%) of all-cause deaths were attributable to COVID-19. Mortality in residents with no direct evidence of infection was 98% higher in care homes with at least one COVID-19 infection or death compared to care homes without (adjusted HR 2.0 [1.6; 2.4]).

**Conclusions:** 1 in 5 residents had symptoms of infection during the pandemic, and excess mortality in LTCFs with outbreaks suggest many deaths occurred in people who were infected with COVID-19, but not tested. Higher occupancy and lower staffing levels increase infection risk. Disease control measures should integrate active surveillance and testing with fun-damental changes in staffing and care home occupancy to protect staff and residents from infection.

@@

Globally the number of COVID-19 cases continues to increase, with substantially higher rates of infection reported in both long-term care facilities (LTCFs) and hospitals.[@PHE2020] In the UK, there are an estimated 400,000 residents living in approximately 11,000 LTCFs for the elderly.[@LaingBuisson2019] Residents of LTCFs are particularly vulnerable to COVID-19 due to their advanced age and high prevalence of comorbidity,[@Zhou2020] and their frequent exposure to infection through close contact with staff members, other residents and contaminated surfaces in the care facility. At the peak of the pandemic, the number of deaths in care home residents was three-fold higher than the equivalent period in in 2019.[@onsDR2020] Staff in LTCFs also have higher aged-standardised rates of COVID-19 related mortality compared to other occupations.[@onsDR2020occ] National statistics suggests two-thirds of excess deaths recorded in residents of LTCFs in the last 6 months involved COVID-19,[@onsDR2020] but this is likely to be an underestimate because many residents were not tested. Understanding the proportion of excess deaths that can be directly and indirectly attributed to COVID-19 infection is important, to fully assess the impact of the pandemic on LTCFs.

Public health strategies to protect the public, residents and staff from COVID-19 include rapid testing, restriction of visitors, and vaccination. Such strategies require knowledge of the burden of and risk factors for infection in residents and staff in LTCFs, linked to outcomes, which draw on evidence from the first wave of the pandemic. Population-based prevalence surveys and studies based on routine data have demonstrated variation in the incidence of infection and case-fatality between countries,[@Fisman2020; @Hollinghurst2020; @Salcher-Konrad2020] but many people with symptoms were not tested, particularly at the start of the pandemic in the context of limited testing capacity. In England, there is no syndromic surveillance for infection in LTCFs in England, and widespread regular testing for SARS-CoV-2 using reverse transcriptase polymerase chain reaction (RT-PCR) was not established for staff and residents in LTCFs until 11 May 2020.[@dhscTesting2020] Prior to this, testing was only available for residents or staff who were admitted to hospital, or as part of Public Health England's (PHE) outbreak investigations which permitted a maximum of five tests per LTCF. Consequently national estimates of incidence and prevalence based on the first wave of infection (February-July 2020) will substantially underestimate the burden of infection in residents and staff in LTCFs.

To our knowledge, there are no studies which have employed population-level active surveillance (daily monitoring to identify possible cases of COVID-19 in residents and staff) in LTCFs to investigate the epidemiology and clinical outcomes of both suspected and confirmed COVID-19 infections. We analysed electronic health records from the Four Seasons Health Care Group, one of the UK’s largest independent provider of residential and nursing care, with the aim of identifying strategies to protect staff and residents in LTCFs from future waves of infection. Our objectives were to estimate incidence of and risk factors for infection, and incidence of mortality in the following groups: (A) residents with no evidence of infection; (B) symptomatic residents; (C) asymptomatic residents with confirmed infection; and (D) symptomatic residents with confirmed infection. We also estimated mortality attributable to COVID-19.

@@

Staff and residents living and working in LTCFs for the elderly between `r format(date_min, "%e %B")` and `r format(date_max, "%e %B %Y")`, which were run by the Four Seasons Healthcare Group (FSHCG) were eligible for study inclusion. The FSHCG provides a combination of residential and nursing care (for residents with medical conditions), which is primarily state funded. Most residents are permanent, but a small proportion receive temporary (respite) care.

In 2020, there were 9,568 beds, representing 9% of all beds in England, Scotland and Northern Ireland (supplementary material 1). 90% of FSHCG LTCFs participated in the whole care home testing programme, implying that all staff and residents were tested for COVID-19 at least once between 11 May and 22 June 2020.
